TBD

### 2018-10-23 manuscript/manuscript.tex
manuscript updates

\subsection*{Quantifying the Fitness Landscape}

To characterize the shape of the fitness landscape in simple and complex food webs, we first used a generalized linear mixed model to quantify selection surfaces on individual traits. We used a binomial error distribution (logit link function) since larval survival (0 or 1) was our response variable and measure of fitness. We specified linear and quadratic terms for each gall trait as well as linear interaction terms between each gall trait as fixed effects in the statistical models. To account for the correlated structure of clutch size (gall level) and female preference (plant level) as well as any independent effects of willow genotype on larval survival, we specified gall ID nested within plant ID nested within plant genotype as random effects. Since we were interested in characterizing the fitness landscape -- the relationship between mean trait values and population mean fitness -- we assumed the mean value of our random effects (i.e. setting them to zero) to estimate selection gradients. Also, the fitness landscape assumes that traits distributions are multivariate normal. To better meet this assumption, we log-transformed clutch size and added a small constant (1) to female preference before log transforming, since our surveys occassionaly estimated zero larval densities. We then scaled all phenotypic traits to mean=0 and SD=1 in order to calculate standardized selection gradients that were comparable across traits and with other studies of natural selection. We used the method of @Janzen1998 to calculate directional ($\beta_{z_i}$), quadratic ($\gamma_{z_i,z_i}$), and correlational ($\gamma_{z_i,z_j}$) selection gradients and used parametric bootstrapping (1000 replicates) to calculate their 95\% confidence intervals (@Bolker2009). We estimated directional selection gradients by excluding quadratic terms and statistical interactions in the model. Note that for visualizing the fitness landscape we restrict trait axes to $\pm 1$ SD of the mean trait value. This emphasizes the fact that we can only reliably estimate the shape of the adaptive landscape near the mean trait values of the population (Arnold 2001, Morrissey 2009).

Rather than imposing selection, parasitoids may themselves influence the expression of herbivore traits. Any influence on trait expression would bias selection gradients acting on those traits. In our system, it was plausible that parasitoids may influence chamber growth by promoting larval feeding (cite), speeding up larva development (cite), or killing larva before they complete their development (cite). Therefore, our estimates of selection on chamber diameter may be positively or negatively biased. To estimate this bias, we subset our data to only include galls where there was variation in larval survival (1 > survival > 0) within the same gall. We then calculated "apparent" selection differentials for each gall by comparing the average chamber diameter of all larva (before "selection") to the average chamber diameter of surviving larva and analyzed separate one-sample t-tests for each food-web treatment. This analysis is based on the assumption that larva within each gall come from the same clutch and therefore should have similar chamber diameters regardless of whether they are parasitized. In general, we found that our estimates of directional selection on chamber diameter were positively biased (Appendix). In other words, our analyses were overestimating the magnitude of selection acting on gall diameter. Therefore, we adjusted our estimates of directional selection on chamber diameter ($\beta_{diam}$) by subtracting the biased selection differentials. Note that selection gradients and selection differentials for chamber diameter were virtually the same (Appendix).

\subsection*{Quantifying Selective Constraints}

The strength and pattern of selective constraints can be measured as the slope and curvature of the fitness landscape (Arnold 1992).

We can translate selection surfaces of individuals to the fitness landscape of a population

To characterize the net effects of food-web complexity on the slope and curvature of *Iteomyia*'s fitness landscape, we took advantage of existing theory that links selection surfaces of individuals to the fitness landscape of the population (Phillips & Arnold 1998, Arnold 2003). Specifically, the slope of the fitness landscape corresponds to the column vector of directional selection gradients:

$$\text{Slope} = \beta = \begin{pmatrix} \beta_{diam} \\ \beta_{clutch} \\ \beta_{pref} \end{pmatrix} $$
whereas the curvature of the fitness landscape is a function of the directional, nonlinear, and correlational selection gradients: 
$$\text{Curvature} = \gamma - \beta \beta^T$$
where \gamma represents the matrix of nonlinear and correlational selection gradients:
$$\gamma = \begin{pmatrix} \gamma_{diam,diam}&& \\ \gamma_{diam,clutch}&\gamma_{clutch,clutch}& \\ \gamma_{diam,pref} & \gamma_{clutch,pref} &\gamma_{pref,pref} \end{pmatrix}$$

Note that we ommitted the upper triangle of the matrix for clarity since it is simply the reflection of the lower triangle. Assuming that there is additive genetic variance and covariance between these traits under selection, then the slope and curvature of the fitness landscape give insight to how the population's mean trait value will change in the next generation as well as how additive genetic variance and covariance changes within a generation.

While making quantitative predictions about trait evolution requires knowledge of the additive genetic variance and covariance of these traits, the slope and curvature of the fitness landscape still give qualitative insight to the evolutionary trajectory of a population.

If we assume that there is additive genetic variance and covariance between these traits, then the matrix describing the curvature of the fitness landscape gives qualitative insight to the selective constraints acting on the population. For example, the diagonal of the curvature matrix dictates (qualitatively) whether the additive genetic variance in each trait will increase ($+$), decrease ($-$), or stay the same ($0$). Similarly, the off-diagonal of the curvature matrix dictates whether selection favors trait integration (positive covariance), a tradeoff (negative covariance), or no change in genetic covariance. In other words, we can get qualitative insight to how food-web complexity influences constraints on the fitness landscape by counting the number of negative sign values along the diagnoal (which imply a decrease in additive genetic variance) and the number of positive or negative signs along the off diagonal (which imply changes in additive genetic covariance that lead to either trait integration or tradeoffs).

All analyses and visualizations were conducted in R (@R2018).

I need to go back to estimate a common alpha coefficient if the treatments do not differ from each other.

If we only retain non-zero selection gradients, then we can

Our two dimensional surfaces are likely representative given the fact that there is no strong evidence of correlational selection (Phillips and Arnold 1989, Evolution)

The curvature ($\gamma - \beta \beta^\text{T}$) of the adaptive landscape describes how natural selection changes additive genetic (co)variances ($\Delta \text{G}$) within a generation. Specifically, the diagonals of the curvature matrix describe the effect of the adaptive landscape on additive genetic variance within a generation; therefore, any negative values along the diagonal are evidence of selective constraints. The off-diagonals of the curvature matrix describe the effect of the adaptive landscape on additive genetic covariances within a generation; therefore, any non-zero values are evidence of selective constraints as they favor trait integration.

\subsection*{Simulating Phenotypic Evolution}

We used the method proposed by Hansen and Houle (2008) to characterize the evolvability of the G-matrices resulting from different adaptive landscapes. Their index of evolvability is theoretically linked to the theory underlying G matrices and adaptive landscapes, and describes the ability of a population to evolve in the direction of selection when stabilizing selection is absent.

% Hansen and Houle 2008
%Alternatively, but not
%inconsistently, evolvability can be viewed in a quantitative
%genetics context as a short-term ability to respond to
%directional selection (Houle, 1992), which depends on
%standing additive genetic variation in the traits under
%selection and on the autonomy of this genetic variation
%from other traits under stabilizing selection (Hansen
%et al., 2003a, b; Hansen, 2003; Hansen & Houle, 2004)

@@

@@

Our key finding was that phenotypic evolution was less constrained for a species that was embedded within a complex food web compared to a simple one. These

@@

@@

%Path relative to the main .tex file 
\graphicspath{ {../analyses/reproduce_analyses_files/figure-html/} }

\begin{figure}[h!]
\includegraphics{Selection-Gradients-1.png}
\caption{Selection gradients acting on \textit{Iteomyia} phenotypes. Points and lines correspond to estimated means and 95\% confidence intervals, respectively. Orange and blue colors correspond to estimated gradients in complex and simple food webs, respectively. Yellow colors correspond to selection gradients imposed by larval parasitoids. Note that we only plot selection gradients for larval parasitoids when they differed from egg parasitoids (simple food web).}
\label{Fig:Gradients}
\end{figure}

\begin{figure}[h!]
\includegraphics{Univariate-Fitness-Landscapes-1.png}
\caption{Adaptive landscape of \textit{Iteomyia} phenotypes in complex vs. simple food webs. Each panel corresponds to a different phenotypic trait: chamber diameter (A); clutch size (B); and female preference (C). Solid lines represent the estimated gradients in complex (orange) and simple (blue) food webs. Transparent lines represent bootstrapped replicates to show the uncertainty in estimated gradients. For clarity, we only display 100 bootstraps even though inferences are based on 1,000 bootstrapped samples. Note the mean larva survival is plotted on a natural log scale to accurately reflect the shape of the adaptive landscape.}
\label{Fig:FL_1D}
\end{figure}

\begin{figure}[h!]
\includegraphics{Multivariate-Fitness-Landscapes-1.png}
\caption{Two dimensional view of adaptive landscapes in complex vs. simple food webs. Each panel corresponds to a different combination of traits: clutch size and chamber diameter (A); clutch size and female preference (B); female preference and chamber diameter (C). Note the mean larva survival is plotted on a natural log scale to accurately reflect the shape of the adaptive landscape.}
\label{Fig:FL_2D}
\end{figure}

>>>>

 Inferring the adaptive landscape from individual selection surfaces assumes that trait distributions are multivariate normal (Lande and Arnold 1983). To approximate this assumption, we log-transformed clutch size and added a small constant (1) to female preference before log transforming, since our surveys occassionaly estimated zero larval densities.  We then scaled all phenotypic traits to mean=0 and SD=1 in order to calculate standardized selection gradients that were comparable across traits and with other studies of natural selection.

Rather than imposing selection, parasitoids may themselves influence the expression of herbivore traits which would bias our measurements of selection. %In our system, it was plausible that parasitoids may influence chamber growth by promoting larval feeding (cite), speeding up larva development (cite), or killing larva before they complete their development (cite). Therefore, our estimates of selection on chamber diameter may be positively or negatively biased. 
To estimate this bias, we subset our data to only include galls where there was variation in larval survival within the same gall (i.e. 1 > mean survival > 0). Using the statistical methods discussed in the next section, we calculated the "apparent" relationship between chamber diameter and larva survival ($\alpha_{\text{Diam}}$). We detected a positive bias ($\alpha_{\text{Diam}}$= 0.35), indicating that unadjusted analyses would overestimate the strength of selection on chamber diameter. To account for this bias, we substracted 0.35 from our estimate of $\alpha_{\text{Diam}}$ with the full dataset prior to calculating directional selection on chamber diameter. Note that this analysis is based on the assumption that larva within each gall come from the same clutch and therefore should have similar chamber diameters regardless of whether they are parasitized. Our data supports this assumption, as we found that gall ID explains 54\% of the variance in chamber diameter.

\subsection*{Quantifying the Adaptive Landscape}

Our analyses consisted of three parts. First, we used generalized linear mixed models (GLMM) to quantify selection surfaces \textemdash linear and nonlinear relationships between absolute fitness ($W$) and phenotypic traits ($z_i$) of individuals \textemdash in each food-web treatment. Second, we translated our approximation of the selection surface into the scale of relative fitness ($w$) in order to calculate selection gradients. Third, we used our estimates of selection gradients to characterize the slope and curvature of the population's adaptive landscape.

\textbf{Selection surface}: Since larva survival was our measure of absolute fitness we used a GLMM that assumed a binomial error distribution (and logit-link function). To approximate the selection surface, we modelled larva survival as a function of food-web treatment as well as linear ($\alpha_{z_i}$), quadratic ($\alpha_{z_i:z_i}$), and linear interactions ($\alpha_{z_i:z_j}$) between each trait. Other approaches have been advocated for approximating the selection surface (Schluter 1988); however, our approach enables us to calculate directional ($\beta_{z_i}$), quadratic ($\gamma_{z_i,z_i}$), and correlational selection ($\gamma_{z_i,z_j}$) gradients, and thus is more appropriate for our approximating the adaptive landscape (cite Arnold). We also allowed these $\alpha$ coefficients to vary between food-web treatments. To account for the nonindependence of individual estimates of clutch size (gall level) and female preference (plant level) as well as any independent effects of willow genotype on larva survival, we specified gall ID nested within plant ID nested within plant genotype as random effects. Including these random effects helps alleviated biased estimates of selection due to environmental covariances between traits and fitness (Rausher 1992 Evolution). With our full statistical model, we used parametric bootstrapping (1,000 replicates) to test whether our food-web treatment altered each $\alpha$ coefficient. If the 95\% confidence interval of this food-web:$\alpha$ effect overlapped with zero, then we removed this statistical interaction and refit the model. Note that our final model of the selection surface still included estimates (and 95\% CI) of our food-web treatment and all possible $\alpha$ coefficients, even if they overlapped with zero, but only a subset of possible food-web:$\alpha$ coefficients. Note also that we estimated linear coefficients ($\alpha_{z_i}$) by excluding quadratic and linear interaction terms in the statistical model (Lande and Arnold 1983). 
%To test how food-web complexity shapes the adaptive landscape, we first used a generalized linear mixed model to quantify selection surfaces on individual traits. We used a binomial error distribution (with logit-link function) since larval survival (0 or 1) was our response variable and measure of fitness. We specified linear and quadratic terms for each gall trait ($z_i$) as well as linear interaction terms between each gall trait as fixed effects ($\alpha$) in the statistical models. These linear ($\alpha_{z_i}$), quadratic ($\alpha_{z_i,z_i}$), and linear interaction ($\alpha_{z_i,z_j}$) terms are key components of our estimates of directional ($\beta_{z_i}$), quadratic ($\gamma_{z_i,z_i}$), and correlational selection ($\gamma_{z_i,z_j}$), respectively. We also specified main effects for food-web treatment and statistical interactions between food-web treatment and each $\alpha$. We used parametric bootstrapping (1,000 replicates) to quantify the effect of each $\alpha$ on larva survival and whether this varied between food-web treatments ($food-web \times \alpha$). If we did not observed an effect of food-web treatment, we fit a reduced model that only include the unique $\alpha$ term. In other words, our statistical model of the adaptive landscape included all possible $\alpha$ terms, but not statistical interactions between food-web treatment and each $\alpha$ term.

\textbf{Selection gradients}: Since we were interested in characterizing the adaptive landscape -- the relationship between mean trait values and population mean fitness -- we assumed the mean value of our random effects (i.e. setting them to zero) when calculating selection gradients. We then used the method of Janzen and Stern (1998) to translate our estimates (and 95\% CI) of each $\alpha$ coefficient into the scale of relative fitness in order to calculate directional ($\beta_{z_i}$), quadratic ($\gamma_{z_i,z_i}$), and correlational ($\gamma_{z_i,z_j}$) selection gradients.

\textbf{Adaptive landscape}: We took advantage of existing theory that translates estimates of selection gradients to the population's adaptive landscape (Phillips and Arnold 1998, Arnold 2003). Specifically, the slope of the adaptive landscape corresponds to the column vector of directional selection gradients:

$$\text{Slope} = \beta = \begin{pmatrix} \beta_{\text{Diam}} \\ \beta_{\text{Clutch}} \\ \beta_{\text{Pref}} \end{pmatrix} $$

whereas the curvature ($C$) of the adaptive landscape is a matrix that is determined by directional, quadratic, and correlational selection gradients:

$$\text{Curvature} = C = \gamma - \beta \beta^T$$

where $\gamma$ represents the matrix of quadratic (diagonal) and correlational (off-diagonal) selection gradients:

$$\gamma = \begin{pmatrix} \gamma_{\text{Diam:Diam}}&& \\ \gamma_{\text{Clutch:Diam}}&\gamma_{\text{Clutch:Clutch}}& \\ \gamma_{\text{Pref:Diam}} & \gamma_{\text{Pref:Clutch}} &\gamma_{\text{Pref:Pref}} \end{pmatrix}$$

Note that we ommitted the upper triangle of the matrix for clarity since it is simply the reflection of the lower triangle. Assuming that there is additive genetic variance and covariance between these traits, then the slope and curvature of the adaptive landscape give insight to changes in the population's mean trait value between generations ($\Delta z = \text{G} \beta$) and G-matrix structure within a generation ($\Delta \text{G} = \text{G} (\gamma - \beta \beta^\text{T}) \text{G}$).

Although quantitative predictions about trait and G-matrix evolution require knowledge of a population's G-matrix, the adaptive landscape can still yield valuable qualitative insight. In particular, we can infer selective constraints acting on a population by inspecting the structure of the matrix describing the curvature of the adaptive landscape. For example, the sign of the values along the diagonal of the curvature matrix dictate whether selection will increase ($+$), decrease ($-$), or cause no change ($0$) in the additive genetic variance of a trait. Similarly, any nonzero covariance terms (off-diagonal) are indicative of trait integration. We propose that you can infer selective constraints acting on a population by counting the number of negative signs along the diagonal (decrease in additive genetic variance) and the number of nonzero terms along the off-diagonal of the matrix (trait integration).

%The curvature ($\gamma - \beta \beta^\text{T}$) of the adaptive landscape describes how natural selection changes additive genetic (co)variances ($\Delta \text{G}$) within a generation. Specifically, the diagonals of the curvature matrix describe the effect of the adaptive landscape on additive genetic variance within a generation; therefore, any negative values along the diagonal are evidence of selective constraints. The off-diagonals of the curvature matrix describe the effect of the adaptive landscape on additive genetic covariances within a generation; therefore, any non-zero values are evidence of selective constraints as they favor trait integration.

\subsection*{Measuring selection on Platygaster sp.}

Our experimental manipulation of food-web complexity enables us to also measure selection on the egg parasitoid (\textit{Platygaster sp.}), albeit with a different approach. Specifically, we can compare the distribution of gall traits that the egg parasitoid is observed in simple vs. complex food webs. In essence, the gall phenotype becomes the extended phenotype of the egg parasitoid, since its attributes can influence the egg parasitoid's survival in the face of intraguild predation (larval parasitoid guild). This cross-sectional analysis allows us to estimate directional and quadratic selection differentials, but not selection gradients.

To quantify directional selection differentials, we used separate linear mixed models with the standardized gall trait as the response variable, food-web treatment as the fixed effect, and the same suite of random effects as for our primary analyses of \textit{Iteomyia} survival (i.e. gall ID, nested within plant ID, nested within plant genotype).

To quantify quadratic selection differentials, we can calculate the difference in variance between simple and complex food webs. We then used Levene's test for homogeneity of variance to determine the statistical significance of these quadratic selection differentials. Note that this result should be interpreted cautiously, since we cannot account for the nonindependence of gall phenotypes due to plant genotype, plant ID, or gall ID.

All analyses and visualizations were conducted in R (@R2018). Note that for visualizing the adaptive landscape we restrict trait axes to $\pm 1$ SD of the mean trait value. This emphasizes the fact that we can only reliably estimate the shape of the adaptive landscape near the mean trait values of the population (Arnold 2001). We also plot mean larva survival on a natural log scale to accurately reflect the shape of the adaptive landscape.

@@

\subsection*{Selection on the extended phenotype of egg parasitoid}

We did not observed any evidence of directional or quadratic selection on chamber diameter ($s_{Diam}=0.06 [-0.14,0.27]; s_{Diam:Diam}=-0.11, P=0.28$) or clutch size ($s_{Clutch}=0.21 [-0.10,0.51]; s_{Clutch:Clutch}=-0.24, P=0.17$) that the egg parasitoid was associated with. However, we did observed directional selection for egg parasitoids to attack larva on high density trees ($s_{Pref}=0.20 [-0.10,0.51]) and an associated decrease in phenotypic variance ($s_{Pref:Pref}=-0.91, P<0.001$).

@@

@@

% If selection was present, it was stronger in complex food webs
% - Difficult to judge how real this is given that it was artificially enhanced due to our food-web treatment.
% - Importance of analyzing components of relative fitness for understanding the mechanisms.

%% 2) Fewer selective constraints in complex vs. simple food webs.

% - Due to fewer traits under directional selection, as well as the presence of disruptive selection.
% - With just 1 less trait under selection, it influenced half of the (co)variance effects on the curvature matrix. 
% - This was a result of conflicting selection pressures in the complex food web.
% - Survival higher at small clutch sizes with egg parasitoids -> I expect this is a density mediated process, where higher local densities of eggs and first-instar larva are easier to detect.
% - Survival higher at large clutch sizes with larval parasitoids -> This could be because neighboring larva could provide a buffer against parasitoid attack, thus enhancing the probability of survival at the individual level.

% - Also due to disruptive selection on female preference.
% - Disruptive pattern appears to be driven by larval guild. This could be a result of them being poorer colonizers and thus reach a saturation effect at higher larval density. This would result in larva at especially high densities to gain a "relative" refuge. In contrast, the egg parasitoid may be an effective forager at the field densities we observed resulting in simply higher parasitism rates. This is consistent with the fact of egg parasitoids being more effective searchers. Appears to be some evidence for this from Amaresekare 2000, Ecology (exploitation-interference tradeoff rather than competition colonization).
% - I think my data is indicative of a type III functional response.

% - Traits vs. ecological variables...female preference and host density as an example.

@@

%Path relative to the main .tex file 
\graphicspath{ {../analyses/} }

\begin{figure}[h!]
\includegraphics[scale=0.5]{complex_simple_foodwebs.jpeg}
\caption{Illustrations of complex (A) and simple (B) food webs associated with the insect herbivore, \emph{Iteomyia salicisverruca}. Black arrows denote the flow of energy in this network of trophic interactions.}
\label{Fig:Concept}
\end{figure}

\begin{figure}[h!]
\includegraphics{reproduce_analyses_files/figure-html/Selection-Gradients-1.png}
\caption{Selection gradients acting on \textit{Iteomyia} phenotypes. Points and lines correspond to estimated means and 95\% confidence intervals, respectively. Orange and blue colors correspond to estimated gradients in complex and simple food webs, respectively. Yellow colors correspond to selection gradients imposed by larval parasitoids. Note that we only plot selection gradients for larval parasitoids when they differed from egg parasitoids (simple food web).}
\label{Fig:Gradients}
\end{figure}

\begin{figure}[h!]
\includegraphics{reproduce_analyses_files/figure-html/Univariate-Fitness-Landscapes-1.png}
\caption{Adaptive landscape of \textit{Iteomyia} phenotypes in complex vs. simple food webs. Each panel corresponds to a different phenotypic trait: chamber diameter (A); clutch size (B); and female preference (C). Solid lines represent the estimated gradients in complex (orange) and simple (blue) food webs. Transparent lines represent bootstrapped replicates to show the uncertainty in estimated gradients. For clarity, we only display 100 bootstraps even though inferences are based on 1,000 bootstrapped samples. Note the mean larva survival is plotted on a natural log scale to accurately reflect the shape of the adaptive landscape.}
\label{Fig:FL_1D}
\end{figure}

\begin{figure}[h!]
\includegraphics{reproduce_analyses_files/figure-html/Multivariate-Fitness-Landscapes-1.png}
\caption{Two dimensional view of adaptive landscapes in complex vs. simple food webs. Each panel corresponds to a different combination of traits: clutch size and chamber diameter (A); clutch size and female preference (B); female preference and chamber diameter (C). Note the mean larva survival is plotted on a natural log scale to accurately reflect the shape of the adaptive landscape.}
\label{Fig:FL_2D}
\end{figure}
